---
title: "Biodiversity in National Parks"
author: "Cindy Te, Jessica Zhang, Shirley Yao, Victoria Limwanich"
date: "11/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
source("scripts/summary_info.R")
source("scripts/aggregate_table.R")
source("scripts/chart_2.R")
```

## Introduction
The purpose and importance of this project is to learn more about the trends of biodiversity. As humans are quickly exceeding one of the nine planetary boundaries, biosphere integrity (especially genetic biodiversity), it is urgent for us find information, analyze it, and understand the trends of biodiversity to develop solutions to prevent biodiversity from further reducing and to help mitigate its negative impacts.

## Summary Information

In a study containing `r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_species"]` species across 
`r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_parks"]` national parks, 
there are `r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_categories"]` different categories of species
that contain a total of `r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_species_type"]` different types of species.
Of these species, there are a total of `r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_rare"]` rare species, 
`r get_summary_info(read.csv("scripts/data/national_parks_biodiversity/species.csv"))["total_endangered"]` of which are endangered species.

## Aggregate Table

This table groups the national parks observations together to summarize the total number of species by category in each park, and then calculate the percentage of species that are of concern, threatened, or endangered.
The table shows that Petrified Forest National Park has the highest percentage (9.26%) of species that are of concern, threatened, or endangered, while Gates Of The Arctic National Park and Preserve has the lowest percentage (1.55%) of species that are of concern, threatened, or endangered.

```{r echo = FALSE}
kable(aggregate_table(read.csv("scripts/data/national_parks_biodiversity/species.csv")), col.names = c(
  "Park",
  "Total of Species",
  "Percent of Species of Concern, Threatened, or Endangered"
))
```

## Chart 1

## Chart 2
```{r echo = FALSE}
chart_2(read.csv("scripts/data/national_parks_biodiversity/parks.csv"), read.csv("scripts/data/national_parks_biodiversity/species.csv"))
```
<br> This chart gives a visualization of the level of biodiversity in different national parks across the United States. It helps give a comparison and ranking of the state of biodiversity across different states/regions. 
This chart was made using the number of unique orders (Carnivora, Rodentia, etc.) in each reported National Park, and then grouping them together by state. From this chart, we can see that the distribution of biodiversity is varied and doesn't have much pattern. The leading state by a large margin, California, has a reported 400 unique orders, and the reported state with the least biodiversity, New Mexico, has 72 unique orders in its national parks. This gives us insight into the amount and state of national parks in the different regions, including the amount of reported orders. 

## Chart 3

